\section*{Conclusion}
\addcontentsline{toc}{section}{Conclusion}

During this internship, we worked first on understanding how $\pi$-calculus worked, and looked at simple typing systems for them. Then we looked specifically at the example shown in~\cite{beffara-2015-unifying} and tried to take that as a base to construct something smaller (no exponential) but that would have a good proof of cut elimination. This is what lead to this work, after many steps of trial and error. This is still not perfect though, as there are many choices made for cut elimination, especially the need to either be able to commute all prefixes with the cut, or the need to be able to commute internal action prefixes with the cut and the cut with itself. There are also many ways to extend this work still, first by extending the system to the additive parts of LL and the $\pi$-calculus, and then adapting the methods to be able to work with the exponential constructs (exponential modalities in MELL and replication in the $\pi$-calculus) as well.

One of the main limitations of this work is the choice to treat the prefixes that project to $\tau$ in the extended system. Although it allows to stay close to the usual calculus, a version with those rules in $\succ$ and $\equiv$ (but with the commutation of action prefixes still left out) would look a lot more like proof nets than this version, and the commutation of $\lambda$ and $\epsilon$ especially would be a reasonable equivalent to the reversibility of $\logpar$ and $\bot$ in the logic part. This version would in turn need some work on the $\pi$-calculus side to allow for commutations of $\pi$ with the parallel, and maybe defining an observational preorder on the $\pi$-calculus at this time.

There is also a limitation in that $\equiv$ is not a bisimulation for $\to$, so this needs to be worked on to either find a good relation and a good congruence to have that, or to write just the right theorem for the confluence of diagrams involving a reduction against a congruence (because $\succ^*$ clearly is too much, but finding the minimal amount of $\succ$ needed for it to work needs research). That is a major difference with the classic calculus, which works up to congruence from the beginning.

The last limitation of importance, that is tied to the first one in the first paragraph, is the behavior of $\leadsto$ and $\succsim$ in regards to the projection, as they do not project to usual known reductions and congruences anymore, but one might be able to define an observational preorder to go around this difficulty. This solution might not be ideal though, research is needed on that as well.

Still, the result is promising, with many ways to improve it and trails to follow for the future, and the properties that emerged from this construct are interesting and welcome for a typing and reduction system on processes. An other thing to study would be the possibility to reverse the projection, that is see if, with a working execution in $\pi$-calculus, we can deduce an annotation that follows this very execution.

At last, the relation of this system to other existing ones is also important, as there is probably a projection of this system's sequents in other existing systems, perhaps $i/o$-types with linearity would be a good lead to start.

\bigskip

\subsection*{Thanks}

I would like to thank my tutor for this internship, Emmanuel Beffara, first, because he was the one who came up with the subject and he was of great help all along, the discussions we had made the work efficient and fast enough to have that much in only 3 months. I want to thank my teachers as well, especially the logic team here, for my first internship here two years ago was what had me interested in mathematical logic in the first place. At last, thank you to all people who had to bear with me during this internship (and in general), notably my friends here, your support was of great help, and to my best friend whose support is so important even though we live so far away.